Sample-Based Determination of Angular Observables

### 2014-06-24 draft.tex
Add and use reference MCSM-2004
Overlooked some indices in Secs. I and II
Fix references and some statements in Sec. III.A

The central tenet of MC integration is the fact that the expectation value $E_P[g]$ of some function
$g(x)$ under the probability density $P(x)$,
\begin{equation}
    E_P[g] \equiv \int g(x) P(x) \dd x
\end{equation}
can be approximated \citeneeded\fred{Robert, Casella} by an MC estimator $\what{E_P[g]}$
\begin{equation}
    \label{eq:mc-id}
    E_P[g] \to \widehat{E_P[g]} \equiv \frac{1}{K} \sum_{k=1} g(x^{(k)})
\end{equation}
due to the strong law of large numbers for $K \to \infty$ assuming
that the variates $x^{(k)}$, $k = 1, \dots, K$ are distributed as
$P$; i.e.,
\begin{equation}
    x^{(k)} \sim P\,.
\end{equation}
Throughout this letter we denote all MC estimators with a wide hat.\\

@@

Application of \refeq{mc-id} then yields \fred{notation is very confusing $P_i$ for $x^{(k)} \sim P$}
\begin{equation}
    \langle P_i\rangle \to \widehat{\langle P_i\rangle} = \frac{1}{K} \sum_{k=1}^{K} f_i(x^{(k)})\,.
\end{equation}
It is often of interest to obtain the $\nuvec$- observables integrated
over certain ranges of $\nuvec$; i.e., binned measurements of $\langle
P_i\rangle$
Therefore, we introduce the bin-integrated quantity
\begin{align}
    \langle P_i\rangle_{\vec{a},\vec{b}}
    & = \int_{\vec{a}}^{\vec{b}} P_i(\nuvec) \dd^M \nu\\
    & = \int_{\vec{a}}^{\vec{b}} \left[\int_{\thvec=0}^{2\pi} f_i(\thvec) P(\nuvec,\thvec) \dd^N\theta\right]\dd^M \nu\\
    & = \int \left[\int_{\thvec=0}^{2\pi} f_i(\thvec) P(\nuvec,\thvec)
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,
        %% \theta^{(m)}(\nuvec - \vec{a}) \theta^{(m)}(\vec{b} - \nuvec)
        \dd^n\theta\right]\dd^m \nu\,,
\end{align}
where the argument of the indicator function $ \mathbf{1}(\vec{a} \le
\nuvec \le \vec{b})$ is to be interpreted componentwise.
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \label{bin-importance}
    \widehat{\langle P_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{N} \sum_{k=1}^{K} f_i(x^{(k)})         \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
This basically reduces the estimation of
$\aver{P_i}_{\vec{a},\vec{b}}$ to a (weighted) counting experiment. In
the limit $N \to \infty$, the random vector $( \dots,
\aver{P_i}_{\vec{a},\vec{b}}, \dots)$ follows a multivariate Gaussian
distribution \citeneeded\fred{Why?  multivariate in $\widehat{P_i}$?}
with a mean vector estimated by \refeq{bin-importance} and covariance
estimated as
\begin{equation}
    \cov[\aver{P_i},\aver{P_j}]_{\vec{a},\vec{b}} \to \what{\cov}[\aver{P}_i, \aver{P}_j]_{\vec{a},\vec{b}}
        = \frac{1}{K - 1} \sum_{k=1}^{K} \big[{f}_i(x^{(k)}) - \widehat{\langle P_i\rangle_{\vec{a},\vec{b}}}\big]\,\big[f_j(x^{(k)}) - \widehat{\langle P_j\rangle_{\vec{a},\vec{b}}}\big]
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
We would like to emphasize that the sample covariance should rapidly \fred{So what value of $K$ is close enought to $\infty$?} converge towards the true covariance matrix.
This will allow to construct constraints on optimized observables with correlated uncertainties
from the experimental constraints on the angular observables. (See \cite{Egede:2008uy,Egede:2010zc,Bobeth:2010wg,Becirevic:2011bp,Bobeth:2012vn,Matias:2012xw,DescotesGenon:2012zf}
for definitions of such optimized observables in $B\to K^*\ell^+\ell^-$ decays).

@@

So far, we assumed that our underlying physical process was accurately described by our PDF $P$.
However, in several interesting processes we might only have an approximative result for $P$.
We will now show that this is the case for the interesting class of four-body decays $B\to P_1 P_2 \ell_1 \bar\ell_2$,
which includes the rare $b\to s$-mediated $B$ decay $B \to K\pi\ell^+\ell^-$ and the $V_{ub}$ suppressed decay
$B\to \pi\pi\ell^+\bar\nu_\ell$. For both examples the PDF $P$ is known in the small-width approximation
and when assuming a pure $P$-wave, resonant final state. An extension to $S-P$ interference has been
studied for $B\to K\pi\ell^+\ell^-$ \cite{Blake:2012xx,Becirevic:2011xx} and for $\bar{B}\to \pi\pi\ell^-\bar\nu_\ell$
\cite{Faller:2013xx}.

@@

We advocate here that is a sensible procedure to perform an expansion in terms of Legendre polynomials $p_{j}^{(m)}$
with respect to the remaining angle $\theta_2$. Using the observables $P_k(\vec{\nu}=(q^2, k^2))$ as defined in \refapp{}
the expansion reads
\begin{equation}
    P_{k}(\vec{\nu},\cos\theta_2) \equiv \sum_{j} \frac{1}{n_{j,m}} P_{k,j}(\vec{\nu}) p_{j}^{(m)}(\cos\theta_2)\sin\theta_2\,,
\end{equation}
where $m=0$ for $n=1,2,6$ and $m=1$ for $n=3,4,5,7,8,9$. The angular observables defined as coefficients of this expansion
have the merit of a well-defined total angular momentum, and thus can be disentangled (see \refapp{partial-waves} for
a derivation). As a consequence of the orthogonality of the Legendre polynomials, mismodelling or lack of modelling of higher
partial-wave observables \emph{does not affect} the method of moments as given in the previous section.
Unfortunately, this benefit on the experimental side is accompanied with a theoretical draw back. Each observable
$P_{k,j}(\vec{\nu})$ consists of an infinite sum of bilinear of partial-wave amplitudes. It remains for theoretical
analyses to estimate or calculate the impact of partial waves beyond the P wave \cite{Das:2014}.\\

>>>>

The central tenet of MC integration is the fact that the expectation value $E_P[g]$ of some function
$g(x)$ under the probability density $P(x)$,
\begin{equation}
    E_P[g] \equiv \int g(x) P(x) \dd x
\end{equation}
can be approximated \cite{MCSM-2004} by an MC estimator $\what{E_P[g]}$
\begin{equation}
    \label{eq:mc-id}
    E_P[g] \to \widehat{E_P[g]} \equiv \frac{1}{K} \sum_{k=1} g(x^{(k)})
\end{equation}
due to the strong law of large numbers for $K \to \infty$ assuming
that the variates $x^{(k)}$, $k = 1, \dots, K$ are distributed as
$P$; i.e.,
\begin{equation}
    x^{(k)} \sim P\,.
\end{equation}
Throughout this letter we denote all MC estimators with a wide hat.\\

@@

Application of \refeq{mc-id} then yields \fred{notation is very confusing $P_i$ for $x^{(k)} \sim P$}
\begin{equation}
    \langle P_i\rangle \to \widehat{\langle P_i\rangle} = \frac{1}{K} \sum_{k=1}^{K} f_i(x^{(k)})\,.
\end{equation}
It is often of interest to obtain the $\nuvec$- observables integrated
over certain ranges of $\nuvec$; i.e., binned measurements of $\langle
P_i\rangle$
Therefore, we introduce the bin-integrated quantity
\begin{align}
    \langle P_i\rangle_{\vec{a},\vec{b}}
    & = \int_{\vec{a}}^{\vec{b}} P_i(\nuvec) \dd^M \nu\\
    & = \int_{\vec{a}}^{\vec{b}} \left[\int_{\thvec=0}^{2\pi} f_i(\thvec) P(\nuvec,\thvec) \dd^N\theta\right]\dd^M \nu\\
    & = \int \left[\int_{\thvec=0}^{2\pi} f_i(\thvec) P(\nuvec,\thvec)
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,
        %% \theta^{(m)}(\nuvec - \vec{a}) \theta^{(m)}(\vec{b} - \nuvec)
        \dd^N\theta\right]\dd^M \nu\,,
\end{align}
where the argument of the indicator function $ \mathbf{1}(\vec{a} \le
\nuvec \le \vec{b})$ is to be interpreted componentwise.
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \label{bin-importance}
    \widehat{\langle P_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{K} \sum_{k=1}^{K} f_i(x^{(k)})         \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
This basically reduces the estimation of
$\aver{P_i}_{\vec{a},\vec{b}}$ to a (weighted) counting experiment. In
the limit $N \to \infty$, the random vector $( \dots,
\aver{P_i}_{\vec{a},\vec{b}}, \dots)$ follows a multivariate Gaussian
distribution \citeneeded\fred{Why?  multivariate in $\widehat{P_i}$?}
with a mean vector estimated by \refeq{bin-importance} and covariance
estimated as
\begin{equation}
    \cov[\aver{P_i},\aver{P_j}]_{\vec{a},\vec{b}} \to \what{\cov}[\aver{P}_i, \aver{P}_j]_{\vec{a},\vec{b}}
        = \frac{1}{K - 1} \sum_{k=1}^{K} \big[{f}_i(x^{(k)}) - \widehat{\langle P_i\rangle_{\vec{a},\vec{b}}}\big]\,\big[f_j(x^{(k)}) - \widehat{\langle P_j\rangle_{\vec{a},\vec{b}}}\big]
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
We would like to emphasize that the sample covariance should rapidly \fred{So what value of $K$ is close enought to $\infty$?} converge towards the true covariance matrix.
This will allow to construct constraints on optimized observables with correlated uncertainties
from the experimental constraints on the angular observables. (See \cite{Egede:2008uy,Egede:2010zc,Bobeth:2010wg,Becirevic:2011bp,Bobeth:2012vn,Matias:2012xw,DescotesGenon:2012zf}
for definitions of such optimized observables in $B\to K^*\ell^+\ell^-$ decays).

@@

So far, we assumed that our underlying physical process was accurately described by our PDF $P$.
However, in several interesting processes we might only have an approximative result for $P$.
We will now show that this is the case for the interesting class of four-body decays $B\to P_1 P_2 \ell_1 \bar\ell_2$,
which includes the rare $b\to s$-mediated $B$ decay $B \to K\pi\ell^+\ell^-$ and the $V_{ub}$ suppressed decay
$B\to \pi\pi\ell^+\bar\nu_\ell$. For both examples the PDF $P$ is known in the small-width approximation
and when assuming a pure $P$-wave, resonant final state. An extension to $S-P$ interference has been
studied for $B\to K\pi\ell^+\ell^-$ \cite{Blake:2012mb,Becirevic:2011bp} and for $\bar{B}\to \pi\pi\ell^-\bar\nu_\ell$
\cite{Faller:2013dwa}.

@@

We advocate here that is a sensible procedure to perform an expansion in terms of Legendre polynomials $p_{j}^{(m)}$
with respect to the remaining angle $\theta_2$. Using the observables $P_k(\vec{\nu}=(q^2, k^2))$ as defined in \refapp{partial-waves}
the expansion reads
\begin{equation}
    P_{k}(\vec{\nu},\cos\theta_2) \equiv \sum_{j} \frac{1}{n_{j,m}} P_{k,j}(\vec{\nu}) p_{j}^{(m)}(\cos\theta_2)\sin\theta_2\,,
\end{equation}
where $m=0$ for $n=1,2,6$ and $m=1$ for $n=3,4,5,7,8,9$. The angular observables defined as coefficients of this expansion
have the merit of a well-defined total angular momentum, and thus can be disentangled for
a derivation). As a consequence of the orthogonality of the Legendre polynomials, mismodelling or lack of modelling of higher
partial-wave observables \emph{does not affect} the method of moments as given in the previous section.
Unfortunately, this benefit on the experimental side is accompanied with a theoretical draw back. Each observable
$P_{k,j}(\vec{\nu})$ consists of an infinite sum of bilinear of partial-wave amplitudes. It remains for theoretical
analyses to estimate or calculate the impact of partial waves beyond the S and P wave contributions. (For $B\to K\pi\ell^+\ell^-$ an initial study has been carried out for contributions up to the D wave \cite{Das:2014}).\\
